---
title: "Advanced Data Visualization"
subtitle: "Going Deeper with R"
output:
  xaringan::moon_reader:
    css: ["style.css", "default"]
    lib_dir: libs
    df_print: paged
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---

```{r child = "setup.Rmd"}
```

```{r eval = TRUE, echo = FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(knitr)
```


---

## Applying best practice to data viz

Highlight findings
Pare down
Focus people's attention
Make it sparkle

1. Use order and color to highlight findings

1. Declutter and simplify

1. Don't make people move their heads

  - Don't turn them to read labels
  - Eliminate legends
  - Add annotations to explain

1. Don't use defaults

1. Combine plots to tell a story

---

## Ordering in plots

Use reorder first, then use factors

---

## Using color to highlight

Do it manually first

Then do it with gghighlight

---


## Don't make people turn their heads

---

## Adding labels directly to plots

Use scales package to show percentages on plots

ggrepel on scatterplot showing association between scores and attendance data

---

## Annotations

Add FT tweet from 1/23/20 from John Murdoch to show it
http://www.thefunctionalart.com/2020/01/an-example-of-how-to-annotate.html

---

## Declutter using themes

Actual themes
- Use hrbrthemes + show various options
- Link to themes RRU blog post

theme() function <- use this as place to demonstrate user-made functions

---

## Using custom fonts

Plotting + saving (i.e. use device = cairo_pdf)

https://www.andrewheiss.com/blog/2017/09/27/working-with-r-cairo-graphics-custom-fonts-and-ggplot/




---


## Unique data viz options in R (aka a tour of ggplot extensions)

- gganimate
- ggridges
- https://github.com/corybrunson/ggalluvial
- [ggalt](https://github.com/hrbrmstr/ggalt) (lollipops, dumbells, etc.)
- [ggbeeswarm](https://github.com/eclarke/ggbeeswarm)
- https://ggraph.data-imaginist.com/
- https://github.com/ricardo-bion/ggradar
- https://github.com/econandrew/ggtreemap
- https://github.com/hrbrmstr/waffle
- Patchwork

Collections
- http://www.ggplot2-exts.org/gallery/
- [From Data to Viz](https://www.data-to-viz.com/)
- https://github.com/erikgahner/awesome-ggplot2




